---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(tidyverse)
library(weathercan)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# headings

<!-- badges: start -->
[![R-CMD-check](https://github.com/paleolimbot/headings/workflows/R-CMD-check/badge.svg)](https://github.com/paleolimbot/headings/actions)
[![Codecov test coverage](https://codecov.io/gh/paleolimbot/headings/branch/master/graph/badge.svg)](https://codecov.io/gh/paleolimbot/headings?branch=master)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of headings is to provide well-tested functions for working with orientation data (headings), including aggregation (e.g., mean, sd) and correction (e.g., declination).

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("paleolimbot/headings")
```
## Example

The headings package can calculate a proper mean and standard deviation of compass headings:

```{r example}
library(headings)

# totally bogus since 0 and 360 are identical!
mean(c(350, 355, 360, 0, 5, 10))
sd(c(350, 355, 360, 0, 5, 10))

# everybody wins!
hdg_mean(c(350, 355, 360, 0, 5, 10))
hdg_sd(c(350, 355, 360, 0, 5, 10))
```

For example, heading aggregation can be used to summarize climate and/or current directions over time.

```{r}
library(tidyverse)
library(weathercan)

# get some climate data
kam <- weathercan::weather_dl(
  station_ids = 51423,
  start = "2016-01-01",
  end = "2016-12-31"
)

kam %>% 
  group_by(month) %>% 
  summarise(
    mean_wtd = hdg_mean(wind_dir, weights = wind_spd, na.rm = TRUE),
    sd_wtd = hdg_sd(wind_dir, weights = wind_spd, na.rm = TRUE)
  )
```

Functions to correct for magnetic declination are also provided for the [IGRF13](https://www.ngdc.noaa.gov/IAGA/vmod/home.html), [WMM2020](https://www.ngdc.noaa.gov/geomag/EMM/index.html), and [EMM2017](https://www.ngdc.noaa.gov/geomag/WMM/DoDWMM.shtml) models.

```{r}
hdg_decl(-64, 45, year = 2021, model = "WMM2020")
hdg_true_from_magnetic(13.40, -64, 45, year = 2021, model = "WMM2020")
hdg_magnetic_from_true(356.51, -64, 45, year = 2021, model = "WMM2020")
```
