---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(tidyverse)
set.seed(39374)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300
)
```

# headings

<!-- badges: start -->
[![R-CMD-check](https://github.com/paleolimbot/headings/workflows/R-CMD-check/badge.svg)](https://github.com/paleolimbot/headings/actions)
[![Codecov test coverage](https://codecov.io/gh/paleolimbot/headings/branch/master/graph/badge.svg)](https://codecov.io/gh/paleolimbot/headings?branch=master)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of headings is to provide well-tested functions for working with orientation data (headings), including aggregation (e.g., mean, sd) and correction (e.g., declination).

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("paleolimbot/headings")
```
## Example

The headings package can calculate a proper mean and standard deviation of compass headings:

```{r example}
library(headings)

# totally bogus since 0 and 360 are identical!
mean(c(350, 355, 360, 0, 5, 10))
sd(c(350, 355, 360, 0, 5, 10))

# everybody wins!
hdg_mean(c(350, 355, 360, 0, 5, 10))
hdg_sd(c(350, 355, 360, 0, 5, 10))
```

For more statistical analyses of directions, you can use the [circular package](https://cran.r-project.org/package=circular). A convenience constructor `hdg_circular()` is provided to construct objects in the way that the circular package expects for compass headings.

Functions to correct for magnetic declination are also provided for the [IGRF13](https://www.ngdc.noaa.gov/IAGA/vmod/home.html), [WMM2020](https://www.ngdc.noaa.gov/geomag/EMM/index.html), and [EMM2017](https://www.ngdc.noaa.gov/geomag/WMM/DoDWMM.shtml) models.

```{r}
hdg_decl(-64, 45, year = 2021)
hdg_true_from_magnetic(13.40, -64, 45, year = 2021)
hdg_magnetic_from_true(356.51, -64, 45, year = 2021)
```

As an example, heading aggregation can be used to summarize climate and/or current directions over time.

```{r}
library(tidyverse)
data("kamloops2016")

kamloops2016 %>% 
  as_tibble() %>%
  filter(is.finite(wind_dir), is.finite(wind_spd)) %>% 
  group_by(month) %>% 
  summarise(
    mean_wtd = hdg_mean(wind_dir, weights = wind_spd),
    sd_wtd = hdg_sd(wind_dir, weights = wind_spd)
  )
```

The headings package also comes with a circle-aware kernel density estimator for visualizing vectors of headings. For circular kernel densities in a non-visualization context, use the [circular package](https://cran.r-project.org/package=circular) (`circular::density.circular()`).

```{r density-base}
plot(
  hdg_density(
    kamloops2016$wind_dir, 
    weights = kamloops2016$wind_spd, 
    na.rm = TRUE
  )
)
```

To use in ggplot2, you will need to extract the values from the density output:

```{r density-ggplot2}
kamloops2016 %>% 
  group_by(month) %>% 
  summarise(
    broom::tidy(hdg_density(wind_dir, weights = wind_spd))
  ) %>% 
  ggplot(aes(x, y)) +
  geom_line() +
  coord_polar() +
  scale_x_continuous(breaks = seq(0, 360, by = 90)) +
  facet_wrap(vars(month)) +
  theme_bw()
```



